---
title: "Coursera Reproducible Research Week2 Proj"
author: "杭城雨"
date: "2020/11/30"
output: html_document
---

## Loading and preprocessing the data
```{r}

download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip", mode="wb")
# unzip data and read 
unzip("activity.zip")
baseData <- read.csv("activity.csv", header = TRUE)
head(baseData)

# change date type to date class 
baseData$date <- as.Date(as.character(baseData$date))
# create NA logical vector
baseDataNA <- is.na(baseData$steps)
# create clean base for later mean calculations
cleanBase <- baseData[!baseDataNA,]
```

##Calculate total number of steps taken each day
```{r }
library(ggplot2)

# aggregate clean non NA steps per day (SUM)
SummedDataByDay <- aggregate(baseData$steps, by=list(baseData$date), sum)
# adjust column names
names(SummedDataByDay)[1] ="date"
names(SummedDataByDay)[2] ="totalsteps"
# top 15 of Summed Steps by day

```

##Make a histogram of the total number of steps taken each day
```{r}
# Plot using ggplot
ggplot(SummedDataByDay, aes(x = totalsteps)) +
  geom_histogram(fill = "steelblue", binwidth=1000) +
  labs(title = "Total Daily Steps", x = "Steps", y = "Frequency")
```

##Calculate and report the mean and median of the total number of steps taken per day
```{r}
# Mean of steps taken per day
mean(SummedDataByDay$totalsteps,na.rm=TRUE)
```
```{r}
#median of steps taken per day
median(SummedDataByDay$totalsteps,na.rm=TRUE)
```
##What is the average daily activity pattern?
#Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
nonNASubset <- baseData[!baseDataNA,]
MeanDataByInterval <- aggregate(nonNASubset$steps, by=list(nonNASubset$interval), mean)
# set the column names
names(MeanDataByInterval)[1] ="interval"
names(MeanDataByInterval)[2] ="steps"

ggplot(MeanDataByInterval, aes(x = interval, y=steps)) +
  labs(title = "Sum of Steps by Interval", x = "interval", y = "steps")+
  geom_line(color="red") 
```
#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxInterval <- MeanDataByInterval[which.max(MeanDataByInterval$steps),]
maxInterval
```
##Imputing missing values
#Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
# generate listing of NA's
missingVals <- sum(baseDataNA)
missingVals
```
#Devise a strategy for filling in all of the missing values in the dataset.
Strategy: Use mean interval steps from Mean Steps for that interval

#Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
# new dataset
baseData2 <- baseData
# NA's in dataset
# dataset minus NA's for Mean calculation
NABase2 <- baseData2[is.na(baseData2$steps),]
cleanBase2 <- baseData2[!is.na(baseData2$steps),]

# generate Mean Data2 by interval
MeanData2ByInterval <- aggregate(cleanBase2$steps, by=list(cleanBase2$interval), sum)
names(MeanData2ByInterval)[1] ="interval"
names(MeanData2ByInterval)[2] ="steps"

# IMPUT METHOD- ATTEMPT 2
baseData2 <- baseData
missingData <- is.na(baseData2$steps)
meanVals <- tapply(cleanBase$steps, cleanBase$interval, mean, na.rm=TRUE, simplify=TRUE)
baseData2$steps[missingData] <- meanVals[as.character(baseData2$interval[missingData])]

# original missing data count
sum(missingData)

# count of NA values
sum(is.na(baseData2$steps))
```

#Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
FullSummedDataByDay <- aggregate(baseData2$steps, by=list(baseData2$date), sum)

names(FullSummedDataByDay)[1] ="date"
names(FullSummedDataByDay)[2] ="totalsteps"
head(FullSummedDataByDay,15)

# Plot using ggplot
ggplot(FullSummedDataByDay, aes(x = totalsteps)) +
  geom_histogram(fill = "steelblue", binwidth=1000) +
  labs(title = "Total Daily Steps", x = "Steps", y = "Frequency")

```

```{r}
# Mean on New Data
mean(FullSummedDataByDay$totalsteps)
# Median on New Data
median(FullSummedDataByDay$totalsteps)
```

#4b) Do these values differ from the estimates from the first part of the assignment?

Yes, the mean is the same but the median has risen 1.19 steps.


#What is the impact of imputing missing data on the estimates of the total daily number of steps?
The effect of using mean data per interval as a data impute method for missing values seems to push overall data towards the mean.

##Are there differences in activity patterns between weekdays and weekends?
```{r}
baseData2$weekday <- weekdays(baseData2$date)
baseData2$weekend <- ifelse (baseData2$weekday == "星期六" | baseData2$weekday == "星期日", "Weekend", "Weekday")

head(baseData2,5)
```

```{r}
MeanDataWeekendWeekday <- aggregate(baseData2$steps, by=list(baseData2$weekend, baseData2$interval), mean)
names(MeanDataWeekendWeekday)[1] ="weekend"
names(MeanDataWeekendWeekday)[2] ="interval"
names(MeanDataWeekendWeekday)[3] ="steps"

ggplot(MeanDataWeekendWeekday, aes(x = interval, y=steps, color=weekend)) +
  geom_line() +
  facet_grid(weekend ~ .) +
  labs(title = "Mean of Steps by Interval", x = "interval", y = "steps")
```


